---
title: "Tutorial 4: Results Object"
author: "Brad Broyles"
format:
  html:
    theme: darkly
    embed-resources: true
editor: visual
---

In this tutorial we will explore the `run_evo3d()` results output object in detail. We will reuse the Hepatitis C Virus E1/E2 example

```{r}
e1 <- system.file("extdata", "e1_hepc_sorted.aln", package = "evo3D")
e2 <- system.file("extdata", "e2_hepc_sorted.aln", package = "evo3D")
pdb <- system.file("extdata", "e1e2_8fsj.pdb", package = "evo3D")
```

```{r}
library(evo3D)

# add some stats #
sc <- list(calc_tajima = TRUE, calc_site_entropy = TRUE)

# adjust some patch parameters #
pc <- list(max_patch = 15, dist_cutoff = NA, distance_method = 'centroid')

results <- run_evo3d(list(e1, e2), pdb, pdb_controls = pc, stat_controls = sc,
                     detail_level = 2, # retain spatial haplotypes #
                     verbose = 0
                     )
```

We have covered `evo3d_df` in detail in tutorial 1.

```{r}
str(results, max.level = 1)
```

### final_msa_subsets

Now `final_msa_subsets` is also populated as we set `detail_level` to 2. Individual spatial haploytpes (structure-informed MSA subsets) can be accessed through `msa_subset_id` column.

```{r}
id = results$evo3d_df$msa_subset_id[1]
results$final_msa_subsets[[id]][1:6, 1:6]
```

### msa_info_sets

Each input MSA will have information stored. The first MSA passed is always called msa1, followed by msa2 and so on. `msa_mat` is the original input data. `ref` has been constructed as a consensus sequence by defualt. `pep` is translated reference sequence. `seq_type` is auto detected nucleotide or protein sequence.

```{r}
str(results$msa_info_sets, max.level = 2)
```

### pdb_info_sets

Each input PDB similar to MSA will have an entry in `pdb_info_sets`. `pdb` is the input object, `chain` indicates the chains in this PDB that were used for downstream MSA↔PDB mappings. `residue_dist` is a distance matrix between residues (default between alpha carbons). `residue_df` is a precursor to `evo3d_df` but contains spatial windows before any mapping to codons. It is worth exploring in a bit more depth.

```{r}
str(results$pdb_info_sets, max.level = 2)
```

```{r}
head(results$pdb_info_sets$pdb1$residue_df)
```

In `residue_df` we see spatial windows (`patch`) as residue indexes before being mapped to codons. Additionally, solvent accessibility information is stored in the `sasa` (absolute accessibility) and `rsa` (relative accessibility) columns.

### aln_info_sets

`aln_info_sets` are stored per PDB entry, gathering all MSA that map to this PDB. It sits between the purely geometric windows of `pdb_info_sets` and the codon indexed (and possible codon collapsed) windows of `evo3d_df`. `coverage` was originally planned for quick coverage plot, but was not implemented as such. `aln_df` is an intermediatate between `residue_df` and the final `evo3d_df`. When homomultimers are in use or multiple PDB files input, `aln_df` stores the original residue anchored but codon indexed spatial windows before these windows are merge into one window per codon. `pos_mat` is likely the most important object here, as it is adjustable and allows `run_evo3d()` to restart upon adjustment.

```{r}
str(results$aln_info_sets, max.level = 2)
```

Examining alignment coverage

```{r}
results$aln_info_sets$pdb1$coverage
```

Each MSA to PDB mapping will have a pos_mat entry. Here showing only the first. It is highly nested, but it seemed optimal to adopt one structure that remained consistent and was easily extractable and inspectable by users.

```{r}
results$aln_info_sets$pdb1$pos_mat$msa1_pdb1_A[1:5,]
```

### call_info

The last element is call_info which stores the parameters of the workflow.

```{r}
str(results$call_info)
```

## Next tutorials:

5.  Custom statistics
